The Concurrent Calculi Formalisation Benchmark

### 2023-10-09 paper/paper.tex
Move discussion after challenge problem description in the paper.

In the rest of this paper we discuss the overall design of the benchmark, describe our three challenge problems, and discuss possible extensions and further work.

\section{Design of the benchmark}
\subsection{Topics we discussed during the meetings}
- Linearity type system dual vs. Vasco's?
- Only linear channels or also unrestricted channels?
- Pi-calculus or CCS for linearity?
- Choice in general
  - Choice can be encoded, so is not necessary
  - Complicates the theory
  - Add as ``future work''
- Choice in scope extrusion
- Extension categories: other topics and small technical differences
- Why not unified syntax?
  - Choice and coinduction does not mix
- Different notions of bisimulation in the coinduction challenge
- Linearity: the context is polymorphic
- Coinduction: why is congruence interesting?
- Why are the challenges not covered by POPLMark/POPLMark Reloaded?
- Scope extrusion does not behave as ``normal'' a binder wrt. semantics, so the usual techniques may not work
- A goal is to build a culture of formalising things in the community
- Linearity is generally interesting
- Encoding techniques are not different from any other field, so there's nothing ``special'' about the formalisations here
- We don't care about the content of messages in coinduction, only synchronization
- Why explicitly bind two endpoints in linearity? To make the duality/polarity connection obvious
- Linearity: Do we care about avoiding deadlocks? No, this is stronger than safety as in well-formedness
- Asynchrony? Not necessary, this is a it more advanced
- Why do we keep the restrictions around in the LTS semantics in scope extrusion?
- ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics

\subsection{Evaluation criteria}

We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
Like previous benchmark authors, we would like to stress that our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types and psi-calculi are not directly covered by our challenge problems.
Most applications will however still need the basic techniques that our challenge problems exercise.
We have designed our challenge problems to exercise the three aspects independently, and they can be solved individually and in any order.
We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.
Solutions to our challenge problems should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges will need to present more elaborate argumentation for their adequacy.

@@

\section{Going beyond the challenge problems}

\section{Conclusion}

@@

\appendix
\section{Challenges}
\input{../challenges/challenges.tex}

>>>>

In the next section we will describe our three challenge problems, before discussing our design choices in \cref{sec:design-discussion}.
We will then discuss possible extensions and further topics in \cref{sec:going-beyond} before concluding in \cref{sec:conclusion}.
The full challenge problems, including proofs, are available in \cref{app:challenges}.

@@

\section{Design of the benchmark}\label{sec:design-discussion}
\subsection{Topics we discussed during the meetings}
- Linearity type system dual vs. Vasco's?
- Only linear channels or also unrestricted channels?
- Pi-calculus or CCS for linearity?
- Choice in general
  - Choice can be encoded, so is not necessary
  - Complicates the theory
  - Add as ``future work''
- Choice in scope extrusion
- Extension categories: other topics and small technical differences
- Why not unified syntax?
  - Choice and coinduction does not mix
- Different notions of bisimulation in the coinduction challenge
- Linearity: the context is polymorphic
- Coinduction: why is congruence interesting?
- Why are the challenges not covered by POPLMark/POPLMark Reloaded?
- Scope extrusion does not behave as ``normal'' a binder wrt. semantics, so the usual techniques may not work
- A goal is to build a culture of formalising things in the community
- Linearity is generally interesting
- Encoding techniques are not different from any other field, so there's nothing ``special'' about the formalisations here
- We don't care about the content of messages in coinduction, only synchronization
- Why explicitly bind two endpoints in linearity? To make the duality/polarity connection obvious
- Linearity: Do we care about avoiding deadlocks? No, this is stronger than safety as in well-formedness
- Asynchrony? Not necessary, this is a it more advanced
- Why do we keep the restrictions around in the LTS semantics in scope extrusion?
- ``Proper'' barbs with a structural congruence in coinduction instead of LTS semantics

\subsection{Evaluation criteria}
We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
Like previous benchmark authors, we would like to stress that our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types and psi-calculi are not directly covered by our challenge problems.
Most applications will however still need the basic techniques that our challenge problems exercise.
We have designed our challenge problems to exercise the three aspects independently, and they can be solved individually and in any order.
We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.
Solutions to our challenge problems should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges will need to present more elaborate argumentation for their adequacy.

\section{Going beyond the challenge problems}\label{sec:going-beyond}

\section{Conclusion}\label{sec:conclusion}

@@

\appendix
\section{Challenges}\label{app:challenges}
\input{../challenges/challenges.tex}
